---
title: "Fit_Linear_Model"
author: "Mikolaj Wieczorek"
date: "10/12/2019"
output:
  md_document:
    variant: markdown_github
---

1. [Get Chevrolet cars in a separate data.frame]
    + [Fit a model]
    + [Create Residual and Predicted variables]
    + [Create a plot of actual vs. predicted values]
    + [Create a plot of residuals vs. predicted]
    + [Create a plot of Price vs. Miles]
    + [Loess fit object]
    + [Create a plot of residuals vs. predicted with Loess]

2. [Form separate data.frames for Impala and Malibu cars]
    + [Fitting a model for Impala and Malibu]
    + [Set up data.frame in order to make predictions for every 10000 miles]
    + [Getting predictions for Impala’s and Malibu‘s vehicles]
    + [Create a plot of both regression lines]
    + [Regression lines intersection point coordinates]

Libraries needed: dplyr, ggplot2, scales.

```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(scales)
```

```{r}
load("~/file_path/UsedCars.RData")
```

```{r eval=FALSE, include=FALSE}
str(UsedCars)
unique(UsedCars$Make)
```
## Get Chevrolet cars in a separate data.frame
Apply a filter: I want only Chevrolets and am going to be only working with Make, Price, and Miles variables.
```{r}
(UsedCars
  %>% select(Make,Price,Miles)
  %>% filter(Make == "Chevrolet")
) -> UsedCars_Chevi
str(UsedCars_Chevi)
```

### Fit a model
```{r}
UsedCars_SLR <- lm(formula = Price ~ Miles, data=UsedCars_Chevi)
summary(UsedCars_SLR)
```

```{r eval=FALSE, include=FALSE}
UsedCars_SLR$residuals
```

### Create Residual and Predicted variables
```{r}
(UsedCars_Chevi %>% 
    mutate(SLR.Residuals = UsedCars_SLR$residuals,
           SLR.Predicted = UsedCars_SLR$fitted.values)) -> Used_Cars_Chevi_Resid.Predict
```

### Create a plot of actual vs. predicted values
Set x-axis limit to start from 0 since price cannot be negative. This will display a warning message: "Removed 18 rows containing missing values (geom_point)."
```{r}
ggplot(Used_Cars_Chevi_Resid.Predict, aes(x = SLR.Predicted, y = Price)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, lwd = 1, col = "blue") +
  xlim(0,25000) +
  labs(x = "Predicted Values", y = "Actual Values", title = "Actual vs. Predicted")
```

### Create a plot of residuals vs. predicted
Outliers estimation at > +/- 2*RMSE
```{r}
#Get RMSE from the summary of fitted model UsedCars_SLR
summary = summary(UsedCars_SLR)
summary$sigma
```
```{r}
#RMSE = 3274
2*3274
```
Thus, outliers will be the points whose residual values are greater than 6548 or smaller than - 6548.
```{r}
ggplot(Used_Cars_Chevi_Resid.Predict, aes(x = SLR.Predicted, y = SLR.Residuals)) +
  geom_point() +
  geom_hline(yintercept = c(0)) +
  geom_hline(yintercept = c(-6548, 6548), lty = "dashed", lwd = 0.3) +
  coord_cartesian(ylim = c(-12000,12000)) +
  xlim(0,20000) +
  scale_y_continuous(labels = comma, breaks = pretty_breaks()) +
  labs(x = "Predicted Values", y = "Residuals", title = "Residuals vs. Predicted")
```

### Create a plot of Price vs. Miles
```{r}
ggplot(Used_Cars_Chevi_Resid.Predict, aes(x = Miles, y = Price)) +
  geom_point() +
  geom_smooth(method = "loess" , se = FALSE) +
  scale_x_continuous(labels = comma, breaks = pretty_breaks()) +
  scale_y_continuous(labels = dollar_format(prefix="$"), breaks = pretty_breaks(n=10)) +
  labs(title = "Price vs. Miles")
  
```


### Loess fit object
```{r}
Chevi_loess <- loess(formula = Price ~ Miles, data = UsedCars_Chevi)
(UsedCars_Chevi
  %>% mutate(Residuals_Loess = Chevi_loess$residuals,
             Predicted_Loess = Chevi_loess$fitted)
) -> Used_Cars_Chevi_Loess_Resid.Predict

```

### Create a plot of residuals vs. predicted with Loess
```{r}
ggplot(Used_Cars_Chevi_Loess_Resid.Predict, aes(x = Predicted_Loess, y = Residuals_Loess)) +
  geom_point() +
  geom_hline(yintercept = c(0)) +
  geom_hline(yintercept = c(-6548, 6548), lty = "dashed", lwd = 0.3) +
  coord_cartesian(ylim = c(-12000,12000)) +
  labs(x = "Predicted Values", y = "Residuals", title = "Residuals vs. Predicted with Loess") +
  scale_x_continuous(labels = comma, breaks = pretty_breaks(n=10)) +
  scale_y_continuous(labels = comma, breaks = pretty_breaks(n=10))
```

Now, I am interested only in Impala and Malibu models. <br></br>
## Form separate data.frames for Impala and Malibu cars
```{r}
( UsedCars
  %>% select(Model,Price,Miles)
  %>% filter(Model == "Impala")
  %>% filter(Miles > 20000 & Miles < 120000)
) -> UsedCars_Impala
```

```{r}
( UsedCars
  %>% select(Model,Price,Miles)
  %>% filter(Model == "Malibu")
  %>% filter(Miles > 20000 & Miles < 120000)
) -> UsedCars_Malibu

```

### Fitting a model for Impala and Malibu
```{r}
UsedCars_Impala_SLR <- lm(formula = Price ~ Miles, data=UsedCars_Impala)
UsedCars_Malibu_SLR <- lm(formula = Price ~ Miles, data=UsedCars_Malibu)
```


### Set up data.frame in order to make predictions for every 10000 miles
```{r}
UsedCars_NewPrediction <- data.frame(Miles = c(20000,30000,40000,50000,60000,70000,80000,90000,100000,110000,120000))
```


### Getting predictions for Impala’s and Malibu‘s vehicles
```{r}
( UsedCars_NewPrediction
  %>% mutate(Impala_Predictions = predict(UsedCars_Impala_SLR,newdata=UsedCars_NewPrediction) )
  %>% mutate(Malibu_Predictions = predict(UsedCars_Malibu_SLR,newdata=UsedCars_NewPrediction) )
) -> UsedCars_NewPrediction
```


### Create a plot of both regression lines
```{r}
ggplot(UsedCars_NewPrediction, mapping=aes(x=Miles,y=Impala_Predictions)) +
  geom_line(mapping=aes(x=Miles,y=Impala_Predictions)) + 
  geom_line(mapping=aes(x=Miles,y=Malibu_Predictions),linetype="dashed")+ 
  labs(x = "Miles", y = "Predictions by Model", title="Predictions by Model vs. Miles") +
  scale_x_continuous(labels = comma, breaks = pretty_breaks(n=10)) +
  scale_y_continuous(labels = dollar_format(prefix="$"), breaks = pretty_breaks(n=10))
```
Dashed: Malibu
Regular: Impala

### Regression lines intersection point coordinates
Find at how many miles Malibu becomes more expensive than Impala. I.g. find the coordinates of a point where the 2 regression lines intersect.

Form a coefficient matrix 'cm' using coefficients from the output of lm() function: UsedCars_Impala_SLR and UsedCars_Malibu_SLR.
```{r}
cm <- rbind(coef(UsedCars_Impala_SLR),coef(UsedCars_Malibu_SLR)) 
c(-solve(cbind(cm[,2],-1)) %*% cm[,1])
```
Plug in the intersection point to check.
```{r}
ggplot(UsedCars_NewPrediction, mapping=aes(x=Miles,y=Impala_Predictions)) +
  geom_line(mapping=aes(x=Miles,y=Impala_Predictions)) + 
  geom_line(mapping=aes(x=Miles,y=Malibu_Predictions),linetype="dashed")+ 
  labs(x = "Miles", y = "Predictions by Model", title="Predictions by Model vs. Miles") +
  scale_x_continuous(labels = comma, breaks = pretty_breaks(n=10)) +
  scale_y_continuous(labels = dollar_format(prefix="$"), breaks = pretty_breaks(n=10)) +
  geom_point(aes(x=89508.494, y=9238.901), colour="blue", size =3) 
```
<br></br>
Malibu is cheaper that Impala when it has less than 89,508 miles. When it has more than that, it becomes more expensive than Impala.




